End-User Software Customization by Direct Manipulation of Tabular Data

### 2020-08-15 paper.tex
finish addressing scrapers

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  In this paper we introduce \emph{table-driven customization}, a new
  way for end users to extend software by direct manipulation without
  doing traditional programming. We augment existing user interfaces
  with a table view showing the structured data inside the application.
  When users edit the table, their changes are reflected in the original
  UI. This simple model accommodates a spreadsheet formula language and
  custom data editing widgets, providing enough power to implement a
  variety of useful extensions.

  We illustrate the approach with Wildcard, a browser extension that
  implements table-driven customization for existing applications using
  web scraping. Through concrete examples, we show that this paradigm
  can support extensions to many real websites, ranging from sorting and
  filtering existing data to adding entire new features. We share
  reflections from our experiences using Wildcard, on both its strengths
  and limitations relative to other customization approaches. Finally,
  we explore how this paradigm might lead to new software architectures
  that encourage this form of end-user customization.
\end{abstract}

@@

We subscribe to MacLean et al's vision of a ``gentle slope''
\citep{maclean1990} free of such ``cliffs,'' where users should only
need to make minimal and incremental investments in skill to achieve
their desired customizations. We seek to contribute to this gentle slope
with a new method for customizing software via direct manipulation,
which requires less skill than scripting but still supports powerful
customizations. We take inspiration from visual database query
interfaces and spreadsheets, which have successfully enabled millions of
end users to compute with data through direct manipulation.

In our proposed paradigm, \emph{table-driven customization}, an
application's UI is augmented with a table view where the user can see
and manipulate the application's internal data. These changes don't just
apply to the table; they also result in immediate changes to the
application's original user interface. The user can sort/filter data in
the UI, inject annotations, pull in related information from other web
services, and more, all using the table as a mediating interface.
Interacting with the table view resembles interacting with a familiar
spreadsheet, but results in customizing an existing application.

\begin{figure*}
\hypertarget{fig:overview}{%
\centering
\includegraphics[width=\textwidth]{media/overview.eps}
\caption{An overview of table-driven customization}\label{fig:overview}
}
\end{figure*}

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement table-driven customization for existing
Web applications. In Section~\ref{sec:example}, we illustrate the end
user experience of table-driven customization through an example
scenario, using Wildcard to add features to Hacker News.

In Section~\ref{sec:architecture}, we explain the system architecture of
table-driven customization. We focus on the \emph{table adapter}
abstraction, which allows many different types of underlying data to be
bidirectionally mapped to a table. We describe several types of table
adapters we've built in Wildcard, and also describe future adapters
supported by the general paradigm.

@@

In Section~\ref{sec:themes}, we discuss some key themes from our work:

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We discuss further how
  table-driven customization relates to a gentle slope of customization.
  We suggest that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain objects rather than low-level representations.
\end{itemize}

Table-driven customization relates to existing work in many areas. Our
goals overlap with many software customization tools, and our methods
overlap with direct manipulation interfaces for working with structured
data, including visual database query systems and spreadsheets. We
explore these connections in Section~\ref{sec:related-work}.

@@

To illustrate the end user experience of table-driven customization, we
consider an example scenario of customizing
\href{https://news.ycombinator.com/}{Hacker News}, a popular tech news
aggregator. Figure~\ref{fig:hacker-news} shows accompanying screenshots.

@@

This formula calls a built-in function that takes a URL as an argument
and uses a public web API to compute an estimated read time for the
URL's contents. The \texttt{link} argument in the formula refers to a
column name in the table; the formula is automatically evaluated across
all rows in the table, using the value of \texttt{link} for each row.

@@

This scenario has shown a few examples of how table-driven
customizations can help a user improve their experience of a website.
Section~\ref{sec:reflections} explains many other use cases, but first
we explain how the system works internally.

@@

Figure~\ref{fig:table-adapter} summarizes the overall architecture of
table-driven customization, using a simplified illustration of the
Hacker News example scenario. The name and points value for each article
is scraped from the web page DOM, and user annotations are loaded from
the brower's local storage.

@@

A key idea in table-driven customization is that a wide variety of data
sources can be mapped to a generic table abstraction. In a relational
database, the table matches the underlying storage format, but in
table-driven customization, the table is merely an \emph{interface
layer}. The data shown in the table is a projection of some underlying
state, and edits to the table can have complex effects on the underlying
state.

Here we describe the abstract interface fulfilled by a table adapter,
and then describe specific types of table adapters we've built so far in
the Wildcard system.

\hypertarget{abstract-interface}{%
\subsubsection{Abstract interface}\label{abstract-interface}}

\emph{Returning a table}: A table adapter exposes a table of data: an
ordered list of records. Each record carries a unique identifier and
associates named attributes with values. Tables have a typed schema, so
the same attributes are shared across all records. We currently support
strings, numeric values, booleans, and datetimes as types. The columns
also carry some additional metadata, such as whether or not they are
read-only or editable.

A table adapter can update the contents of a table at any time in
response to changes in the underlying state (e.g., a DOM scraping
adapter can update the table when the page body changes). When data
changes, the query view is reactively updated in response.

\emph{Handling edits}: The query engine can issue a request to a table
adapter to make an edit to a record. The meaning of making an edit can
vary depending on the adapter: in the local storage adapter, a new value
may be persisted into local storage; in the DOM scraping adapter, an
edit may result in changing the value of a form field.

In addition, the query engine also sends additional information about
the combined query view to each table adapter:

\emph{Sorting/filtering}: When the user sorts or filters the query view,
an ordered list of visible IDs is sent to each table adapter. The DOM
scraping adapter uses this information to change the list of rows shown
in the web page.

\emph{Data from other tables}: The query engine provides each table
adapter with the entire combined table shown to the user. The DOM
scraping adapter uses this for injecting annotations---values from other
tables are added to the original web page.

\emph{Currently selected record}: As the user clicks around the table
view, the query engine broadcasts the record currently selected by the
user to each table adapter. The DOM scraping adapter uses this
information to highlight the row in the page that corresponds to the
selected row in the table, which helps the user understand the mapping
between the table and the original UI.

@@

DOM scraping adapters are the essential component that enables Wildcard
to interface with an existing website UI. In addition to the standard
web scraping problem of extracting a table of data from the DOM, a
scraping adapter must also manipulate the DOM to reorder rows, edit form
entries, and inject annotations as the table is edited.

Because every website has unique content, we rely on programmers to
create a DOM scraping adapter for each individual website to make it
available for customization in Wildcard. To make this approach viable,
we have built a generic DOM scraping adapter which a programmer can
configure with the minimal site-specific parts.

\emph{Basic metadata}: The programmer specifies URL patterns that the
adapter should activate on, and a set of column names and types to
return in the table.

\emph{Specifying scraping logic}: The main component of a site-specific
adapter is a single Javascript function that extracts relevant elements
from the page using web scraping technqiues. For each scraped row of
data, the programmer must return a unique identifier, and values for
each column in the table.

The generic adapter then wraps this scraping function to implement the
table adapter interface. For example, when the user sorts the table
view, the generic adapter synchronizes the sort order to the DOM by
removing the DOM elements corresponding to the rows in the table and
reinserting them back into the page in the new order.

\emph{Choosing an ID}: When possible, it is best to choose a server-side
identifier that remains stable across pageloads. This enables user
annotations persisted in local storage to be associated with the same
records on subsequent pageloads. We have found that it's usually
possible to find such an identifier; for example, each item in a page
often contains a link to a page with more details, with a URL that
contains a stable ID.

@@

The default option is to return a DOM element, in which case the generic
adapter extracts the text contents of the DOM element and casts them to
the type of the column. For example, if the scraping function returns a
\texttt{\textless{}div\textgreater{}} containing the string ``1.23'',
and the column has a numeric type, the generic adapter extracts the
number 1.23. The advantage of returning a DOM element is that the value
is editable---when the user changes the value in the table, the generic
adapter can simply overwrite the inner contents of the DOM element. In
practice, this is most useful for editing the value of form input
fields. It's technically possible to edit the contents of any element in
the page (e.g.~editing the title of a post on Hacker News), but this
often isn't particularly useful, so adapter authors have the ability to
mark fields as readonly.

@@

\emph{Other site-specific customization}: The programmer can optionally
override other behavior of the generic DOM scraping adapter, including
logic for when to re-run the scraping function in response to page
changes, how to style injected annotations, and how to style the
currently selected row in the table. We've chosen sensible defaults that
generally result in a workable adapter without needing to change these
options, but they can improve the experience on a specific website. For
example, by default we highlight the currently selected row with a blue
border, but on some sites this is hard to see or breaks the layout, so
setting a background color works better.

@@

\hypertarget{future-adapters}{%
\subsubsection{Future Adapters}\label{future-adapters}}

We have designed the table adapter API to be general enough to support
other types of useful adapters in the future. Here are two examples:

\textbf{Integrated website adapters}: A key benefit of the table adapter
abstraction is that Wildcard is not coupled to web scraping as the only
means for integrating with existing sites, but can also accommodate
first party developers adding support directly into their own websites.
An ``integrated website adapter'' installed by the developer could
directly accesses the internal state of the application, providing the
same functionality as a DOM scraping adapter but in a more robust way.

With the advent of rich frontend web frameworks, structured application
state is now often available in the web client. We suspect it is
possible to create plugins for frontend frameworks that expose this
state to Wildcard with only minimal effort from the application
developers. To test this hypothesis, we created an early
proof-of-concept of an integrated website adapter for the popular
\href{https://redux.js.org/}{Redux} Javascript state management library.
In Redux, app developers conveniently already represent the state of a
user interface as a single object, so to expose a table of data to
Wildcard, the developer simply writes a function that maps the
centralized state object to a table.

Customizability is sometimes a key selling point for software. An
integrated website adapter would provide a way for developers to
integrate with an ecosystem of formulas and customization tools without
needing to build all that functionality from scratch.

\textbf{Shared storage adapter}: It would be useful to share user
annotations among users and across devices---for example,
collaboratively taking notes with friends on options for places to stay
together on a trip. The existing local storage adapter could be extended
to share live synchronized data with other users. This could be achieved
through a centralized web server, or through P2P connections that might
provide stronger privacy guarantees.

@@

\begin{table*}[]
\begin{tabular}{llrl}
\hline
\textbf{Website} & \textbf{Description} & \textbf{LOC} & \textbf{Example customizations}                                                              \\ \hline
Airbnb           & Travel               & 73                                       & Add Walkability Scores to listings. Sort listings by price.                           \\
Amazon           & Online shopping      & 99                                       & Sort third party sellers by total price, including fees.                                \\
Blogger          & Blogging             & 36                                       & Use alternate text editor to edit blog posts.                                                 \\
Expedia          & Travel               & 41                                       & Use alternate datepicker to enter travel dates.                                               \\
Flux             & Data portal          & 67                                       & Use Wildcard as a faster table editor for editing lab results.                                \\
Github           & Code repository      & 62                                       & Sort a user's code repositories by stars to find popular work.                                \\
Hacker News      & News                 & 69                                       & Add read times to links. Filter out links that have been read. \\
Instacart        & Grocery delivery     & 48                                       & Sort groceries by price and category. Take notes on items.                                   \\
Uber Eats        & Food delivery        & 117                                      & Sort/filter restaurants by estimated delivery ETA and price.                                 \\
Weather.com  & Weather              & 51                                       & Sort/filter hourly weather to find nice times of day.                                        \\
Youtube          & Videos               & 80                                       & Sort/filter videos by length, to find short videos to watch.                                 \\ \hline
\end{tabular}
\caption{A list of table-driven customizations that we have implemented using Wildcard.}
\label{tab:websites}
\end{table*}

To evaluate table-driven customization in practice, we built the
Wildcard browser extension, which implements table-driven customization
in the context of existing websites. It is implemented in Typescript,
and works across three major browsers: Chrome, Firefox, and Edge.

@@

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate aspects of using the system in practice.

@@

There are many customizations that are not possible to implement with
table-driven customization. Some of the limitations are specific to the
current implementation of the Wildcard extension, but others are more
fundamental to the general paradigm.

@@

Our second evaluation area relates less to the conceptual approach of
table-driven customization, and more to the specific implementation of
customizing existing web applications. In order for third-party
customization through Wildcard to succeed, it is important that creating
usable adapters for existing websites takes minimal effort.

@@

\hypertarget{sec:themes}{%
\section{Key themes}\label{sec:themes}}

Here we discuss some themes that we've explored with this work, which go
beyond our specific approach and address broader issues in software
customization.

@@

Imperative programming is a reasonable choice as the model for building
customizations. Turing-complete programming provides a high ceiling for
possible customizations, and a sequence of commands is a natural fit for
automations that simulate a series of steps taken by the user. There is,
however, a serious drawback to this approach. MacLean et al
\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users wanting to customize their applications to learn
programming creates an abrupt ``cliff,'' exacting a significant
investment in learning even to implement the simplest customizations.
Another goal of MacLean et al is to make it ``as easy to change the
environment as it is to use it''---at least for some subset of changes.
But in scripting languages, the experience of customization does not
remotely resemble the experience of use.

>>>>

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}
%%
%% end of the preamble, start of the body of the document source.

\usepackage{listings}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\begin{document}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  In this paper we introduce \emph{data-driven customization}, a new way
  for end users to extend software by direct manipulation without doing
  traditional programming. We augment existing user interfaces with a
  table view showing the structured data inside the application. When
  users edit the table, their changes are reflected in the original UI.
  This simple model accommodates a spreadsheet formula language and
  custom data editing widgets, providing enough power to implement a
  variety of useful extensions.

  We illustrate the approach with Wildcard, a browser extension that
  implements data-driven customization for existing applications using
  web scraping. Through concrete examples, we show that this paradigm
  can support extensions to many real websites, ranging from sorting and
  filtering existing data to adding entire new features. We share
  reflections from our experiences using Wildcard, on both its strengths
  and limitations relative to other customization approaches. Finally,
  we explore how this paradigm might lead to new software architectures
  that encourage this form of end-user customization.
\end{abstract}

@@

We subscribe to MacLean et al.'s vision of a ``gentle slope''
\citep{maclean1990} free of such ``cliffs,'' where users should only
need to make minimal and incremental investments in skill to achieve
their desired customizations. We seek to contribute to this gentle slope
with a new method for customizing software via direct manipulation,
taking inspiration from visual database query interfaces and
spreadsheets, which have successfully enabled millions of end users to
compute with data through direct manipulation.

In our proposed paradigm, \emph{data-driven customization}, an
application's UI is augmented with a table view where the user can see
and manipulate the application's internal data. These changes don't just
apply to the table; they also result in immediate changes to the
application's original user interface. The user can sort/filter data in
the UI, inject annotations, pull in related information from other web
services, and more, all using the table as a mediating interface.
Interacting with the table view resembles interacting with a familiar
spreadsheet, but results in customizing an existing application.

\begin{figure*}
\hypertarget{fig:overview}{%
\centering
\includegraphics[width=\textwidth]{media/overview.eps}
\caption{An overview of data-driven customization}\label{fig:overview}
}
\end{figure*}

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement data-driven customization for existing
Web applications. In Section~\ref{sec:example}, we concretely
demonstrate the end user experience of using Wildcard to add features to
a news website.

In Section~\ref{sec:architecture}, we explain the system architecture of
data-driven customization. We focus on the \emph{table adapter}
abstraction, which allows many different types of underlying data to be
bidirectionally mapped to a table. We describe several types of table
adapters we've built in Wildcard, and also describe future adapters
supported by the general paradigm.

@@

In Section~\ref{sec:visions}, we discuss some key themes from our work:

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We discuss further how
  data-driven customization relates to a gentle slope of customization.
  We suggest that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain data rather than low-level representations.
\end{itemize}

data-driven customization relates to existing work in many areas. Our
goals overlap with many software customization tools, and our methods
overlap with direct manipulation interfaces for working with structured
data, including visual database query systems and spreadsheets. We
explore these connections in Section~\ref{sec:related-work}.

@@

To concretely illustrate the user experience of data-driven
customization, we present a scenario of customizing
\href{https://news.ycombinator.com/}{Hacker News}, a popular tech news
aggregator. Figure~\ref{fig:hacker-news} shows accompanying screenshots.

@@

This formula calls a built-in function \texttt{ReadTimeInSeconds} that
uses a third-party public web API to compute an estimated read time for
the URL's contents. The \texttt{link} argument in the formula refers to
a column name in the table; the formula is automatically evaluated
across all rows in the table, using the value of \texttt{link} for each
row.

@@

This scenario has shown a few examples of how data-driven customizations
can help a user improve their experience of a website.
Section~\ref{sec:reflections} explains many other use cases and contexts
where the technique applies, but first we explain how the system works
internally.

@@

Figure~\ref{fig:table-adapter} summarizes the overall architecture of
data-driven customization, using a simplified illustration of the Hacker
News example scenario. The name and points value for each article is
scraped from the web page DOM, and user annotations are loaded from the
brower's local storage.

@@

A key idea in data-driven customization is that a wide variety of data
sources can be mapped to a generic table abstraction. In a relational
database, the table matches the underlying storage format, but in
data-driven customization, the table is merely an \emph{interface
layer}. The data shown in the table is a projection of some underlying
state, and edits to the table can have complex effects on the underlying
state.

Here we describe the three types of table adapters we have built so far
in Wildcard. These do not represent an exhaustive set of all possible
table adapters---in Section~\ref{sec:visions} we discuss other types of
adapters that would fit well into the general paradigm.

@@

DOM scraping adapters enable Wildcard to interface with an existing
website UI. A DOM scraping adapter fulfills the standard web scraping
task of extracting a table of data from the DOM, but it must also handle
the reverse direction: manipulating the DOM to reflect edits to the
table.

Currently, we require DOM scraping adapters to be programmed manually
for each website using Javascript code. It might seem that this
prohibits non-programmer users from using the system, but we solve this
problem via a shared repository of adapters. Once an adapter is
programmed for a website, it is added to a shared code repository,
enabling any end user to perform customizations on that website.

There are other ways a DOM scraping adapter could be produced which
would reduce this dependence on programmers: an end user could specify
the scraping logic via demonstration, or the desired data table could be
automatically inferred from the page. While we are interested in these
techniques and discuss them later in Section~\ref{sec:visions}, we
believe that a shared repository of manually programmed adapters is a
pragmatic starting point; given that many users visit the same popular
websites,a critical mass of adapters could serve the needs of many
users.

To ease the burden on programmers of creating these adapters, we have
created a framework makes the process feel more like writing
unidirectional scraping code than performing a complex bidirectional
synchronization. The key idea is this: programmers return pointers to
DOM elements representing table rows and table cells; Wildcard extracts
data from these DOM elements, but it also uses the pointers to
synchronize table edits back into the page. For example, when the user
sorts the table, the DOM elements representing the table rows are moved
around in the DOM to reflect the new sorted order.

\begin{figure*}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{const}\NormalTok{ HNAdapter }\OperatorTok{=} \AttributeTok{createDomScrapingAdapter}\NormalTok{(}\OperatorTok{\{}
  \DataTypeTok{name}\OperatorTok{:} \StringTok{"Hacker News"}\OperatorTok{,}

  \CommentTok{// Specify when the adapter should be enabled, based on current URL}
  \AttributeTok{enabled}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return}\NormalTok{ (}
      \AttributeTok{urlExact}\NormalTok{(}\StringTok{"news.ycombinator.com/"}\NormalTok{) }\OperatorTok{||}
      \AttributeTok{urlContains}\NormalTok{(}\StringTok{"news.ycombinator.com/news"}\NormalTok{) }\OperatorTok{||}
      \AttributeTok{urlContains}\NormalTok{(}\StringTok{"news.ycombinator.com/newest"}\NormalTok{)}
\NormalTok{    )}\OperatorTok{;}
  \OperatorTok{\},}

  \CommentTok{// Define the name and type of each column in the table}
  \DataTypeTok{attributes}\OperatorTok{:}\NormalTok{ [}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"id"}\OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,} \DataTypeTok{hidden}\OperatorTok{:} \KeywordTok{true} \OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"rank"}\OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"numeric"} \OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"title"}\OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"} \OperatorTok{\},}
    \OperatorTok{\{} \DataTypeTok{name}\OperatorTok{:} \StringTok{"link"}\OperatorTok{,} \DataTypeTok{type}\OperatorTok{:} \StringTok{"text"} \OperatorTok{\},}
    \CommentTok{// ... other columns omitted for brevity}
\NormalTok{  ]}\OperatorTok{,}

  \CommentTok{// Iterate over DOM elements, returning information about each row}
  \AttributeTok{scrapePage}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return} \VariableTok{Array}\NormalTok{.}\AttributeTok{from}\NormalTok{(}\VariableTok{document}\NormalTok{.}\AttributeTok{querySelectorAll}\NormalTok{(}\StringTok{"tr.athing"}\NormalTok{)).}\AttributeTok{map}\NormalTok{((el) }\KeywordTok{=>} \OperatorTok{\{}
      \KeywordTok{let}\NormalTok{ detailsRow }\OperatorTok{=} \VariableTok{el}\NormalTok{.}\AttributeTok{nextElementSibling}\OperatorTok{;}
      \KeywordTok{let}\NormalTok{ spacerRow }\OperatorTok{=} \VariableTok{detailsRow}\NormalTok{.}\AttributeTok{nextElementSibling}\OperatorTok{;}

      \ControlFlowTok{return} \OperatorTok{\{}
        \CommentTok{// Return a unique ID for each row}
        \DataTypeTok{id}\OperatorTok{:} \AttributeTok{String}\NormalTok{(}\VariableTok{el}\NormalTok{.}\AttributeTok{getAttribute}\NormalTok{(}\StringTok{"id"}\NormalTok{))}\OperatorTok{,}

        \CommentTok{// Return DOM elements corresponding to this row}
        \CommentTok{// (this enables moving/hiding the elements for sorting/filtering)}
        \DataTypeTok{rowElements}\OperatorTok{:}\NormalTok{ [el}\OperatorTok{,}\NormalTok{ detailsRow}\OperatorTok{,}\NormalTok{ spacerRow]}\OperatorTok{,}

        \CommentTok{// Return data for each column}
        \DataTypeTok{dataValues}\OperatorTok{:} \OperatorTok{\{}
          \DataTypeTok{rank}\OperatorTok{:} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{"span.rank"}\NormalTok{)}\OperatorTok{,}
          \DataTypeTok{title}\OperatorTok{:} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{"a.storylink"}\NormalTok{)}\OperatorTok{,}
          \DataTypeTok{link}\OperatorTok{:} \VariableTok{el}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{"a.storylink"}\NormalTok{).}\AttributeTok{getAttribute}\NormalTok{(}\StringTok{"href"}\NormalTok{)}\OperatorTok{,}
          \CommentTok{// ... other columns omitted for brevity}
        \OperatorTok{\},}

        \CommentTok{// Specify where annotations should be injected, and what they should look like}
        \DataTypeTok{annotationContainer}\OperatorTok{:} \VariableTok{detailsRow}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{"td.subtext"}\NormalTok{) }\ImportTok{as}\NormalTok{ HTMLElement}\OperatorTok{,}
        \DataTypeTok{annotationTemplate}\OperatorTok{:} \VerbatimStringTok{\textasciigrave{}| <span style="color: \#f60;">$annotation</span>\textasciigrave{}}\OperatorTok{,}
      \OperatorTok{\};}
    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
  \OperatorTok{\},}
\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

\caption{Source code for the Hacker News scraper. Some details removed for brevity.}\label{fig:scraper-code}
\end{figure*}

Figure~\ref{fig:scraper-code} shows an example of the scraper code used
for the Hacker News example (with some code eliminated for brevity.) It
defines the following main components:

\begin{itemize}
\tightlist
\item
  \texttt{enabled}: defines when this adapter should run, usually based
  on the active URL in the browser.\footnote{Currently Wildcard can only
    show a single table at a time, so if multiple adapters are enabled
    for a single page, we arbitrarily pick one. It would be a
    straightforward extension to allow the user to switch between
    multiple possible tables available on the page.}
\item
  \texttt{attributes}: defines a schema for the table, with a name and
  type for each column
\item
  \texttt{scrapePage}: defines a scraping function which returns an
  array of objects, each containing the data for a single row of the
  table.
\end{itemize}

Here are some of the concerns that emerge when building adapters in
practice:

\emph{Choosing a row ID}: When possible, it is best to choose a
server-side identifier that remains stable across pageloads. This
enables user annotations persisted in local storage to be associated
with the same records on subsequent pageloads. We have found that it's
usually possible to find such an identifier; for example, each item in a
page often contains a link to a page with more details, with a URL that
contains a stable ID.

@@

The default option is to return a DOM element, in which case the generic
adapter extracts the text contents of the DOM element and casts them to
the type of the column. The advantage of returning a DOM element is that
the value is editable---when the user changes the value in the table,
the generic adapter can simply overwrite the inner contents of the DOM
element.

@@

\emph{Optional overrides}: In order to turn a unidirectional scraping
function into a bidirectional scraping adapter, there are a number of
behaviors that must be specified:

\begin{itemize}
\tightlist
\item
  when should the scraping function be re-run in response to changes on
  the page?
\item
  how should injected annotations appear in table rows?
\item
  when the user selects a row in the table, how should the corresponding
  row in the DOM be highlighted?
\end{itemize}

The scraping framework defines sensible defaults that work well on many
sites, but the programmer can optionally override them to provide better
site-specific behavior. For example, the Hacker News adapter specifies
annotation options which make user annotations appear more naturally in
the context of the original design.

@@

@@

\begin{table*}[]
\begin{tabular}{llrl}
\hline
\textbf{Website} & \textbf{Description} & \textbf{LOC} & \textbf{Example customizations}                                                              \\ \hline
Airbnb           & Travel               & 73                                       & Add Walkability Scores to listings. Sort listings by price.                           \\
Amazon           & Online shopping      & 99                                       & Sort third party sellers by total price, including fees.                                \\
Blogger          & Blogging             & 36                                       & Use alternate text editor to edit blog posts.                                                 \\
Expedia          & Travel               & 41                                       & Use alternate datepicker to enter travel dates.                                               \\
Flux             & Data portal          & 67                                       & Use Wildcard as a faster table editor for editing lab results.                                \\
Github           & Code repository      & 62                                       & Sort a user's code repositories by stars to find popular work.                                \\
Hacker News      & News                 & 69                                       & Add read times to links. Filter out links that have been read. \\
Instacart        & Grocery delivery     & 48                                       & Sort groceries by price and category. Take notes on items.                                   \\
Uber Eats        & Food delivery        & 117                                      & Sort/filter restaurants by estimated delivery ETA and price.                                 \\
Weather.com  & Weather              & 51                                       & Sort/filter hourly weather to find nice times of day.                                        \\
Youtube          & Videos               & 80                                       & Sort/filter videos by length, to find short videos to watch.                                 \\ \hline
\end{tabular}
\caption{A list of data-driven customizations that we have implemented using Wildcard.}
\label{tab:websites}
\end{table*}

To evaluate data-driven customization in practice, we built the Wildcard
browser extension, which implements data-driven customization in the
context of existing websites. It is implemented in Typescript, and works
across three major browsers: Chrome, Firefox, and Edge.

@@

We have found that data-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate aspects of using the system in practice.

@@

There are many customizations that are not possible to implement with
data-driven customization. Some of the limitations are specific to the
current implementation of the Wildcard extension, but others are more
fundamental to the general paradigm.

@@

Our second evaluation area relates less to the conceptual approach of
data-driven customization, and more to the specific implementation of
customizing existing web applications. In order for third-party
customization through Wildcard to succeed, it is important that creating
usable adapters for existing websites takes minimal effort.

@@

\hypertarget{sec:visions}{%
\section{Vision}\label{sec:visions}}

\begin{itemize}
\tightlist
\item
  Wildcard just one instance of a broader data-driven customization

  \begin{itemize}
  \tightlist
  \item
    particular view: table
  \item
    particular context: web
  \end{itemize}
\item
  here we share some of the underlying ideas. go beyond Wildcard, room
  to grow
\end{itemize}

@@

Imperative programming is a reasonable choice as the model for building
customizations. Turing-complete programming provides a high ceiling for
possible customizations, and a sequence of commands is a natural fit for
automations that simulate a series of steps taken by the user. There is,
however, a serious drawback to this approach. MacLean et
al.~\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users wanting to customize their applications to learn
programming creates an abrupt ``cliff,'' exacting a significant
investment in learning even to implement the simplest customizations.
Another goal of MacLean et al.~is to make it ``as easy to change the
environment as it is to use it''---at least for some subset of changes.
But in scripting languages, the experience of customization does not
remotely resemble the experience of use.
